Groth16 zk-SNARK Proof Verification on Freeton : Three Use Cases

### 2021-07-14 submission/submission.tex
update submission with Euler and Sudoku parts

\begin{description}
	\item[Code:] All our code comes with a README and tests. This document is provided as an overview of our use cases for \zksnarks{} on the Freeton blockchain but you can also directly plunge into the code. All our code is under the GPL license and is hosted at \url{https://github.com/OCamlPro/devex-18-zk-contest}.

@@

	\item[Keys on contracts: ] For all of our smart contracts, we provided, along with the verification key as needed by the \verb|vergroth16| instruction, the proving key (in a zipped) on the smart contract. This enables users to easily find it in order to generate their proofs. In case the contract deployer should decide to keep the possibility of changing the verification and proving keys of the problem(s), this also makes it easiest to follow changes before submitting a solution.

@@

\section{A toy example: Sudoku}
\label{section_sudoku}

@@

\emph{Note: You may skip this section if you don't want to focus too much on the details on a first reading.}

@@

  TODO: mettre un exemple
  copier coller la commande

	\begin{itemize}
		\item To create the original proving and verification keys. This step is independent of any given Sudoku instance and should only be done once.
		\item To generate a proof given an instance and a solution. The instance is given as a text file such as

@@

		Note that if your solution is not correct, this step will fail.
	\end{itemize}
    \item[The Sudoku smart contract:] The main function of this smart contract is
    \begin{lstlisting}[language=Solidity]
    function submit(bytes proof)
        public view returns (bool res, string blob_str)
    \end{lstlisting}
    It will check the input proof against the verification key
    \lstinline[language=Solidity]|v_key| and the current instance
    \lstinline[language=Solidity]|m_instance|, which are both local to
    the contract and set in the constructor. This instance is stored
    as an array of \lstinline[language=Solidity]|fixed_value|:
    \begin{lstlisting}[language=Solidity]
    struct fixed_value {
      uint8 i;
      uint8 j;
      uint8 value;
    }

@@

The previous example is somewhat limited by the fact that Sudokus are easily solvable problems, either by hand or with a computer; they are not a realistic use case of \zksnarks{}\footnote{Although they hint at a possible use for \zksnarks{}: checking a result without implementing its full logic in Solidity may save precious storage and computations on the blockchain.}.

@@

\subsection{Brute-force protection}
\label{subsec:brute-force}

@@

\begin{description}
	\item[The cli:]
	\item[The EulerRoot TIP3-style contract:] Following the TIP3 conventions, this is the root contract to two types of leaves: EulerUser and EulerProblem. The EulerUser contract represents one user and the list of problems she has solved. The EulerProblem contract represents one problem and contains, among other things, the top ten leaderboard of the fastest solvers.
\end{description}

@@

This use case includes 3 different components:
\begin{description}
\item[PinCode Client:] it is a C++ program linked to the Blueprint
  \zksnarks{} library. It can be used in two different ways:
  \begin{itemize}
  \item To create the initial circuit using the passphrase. The client
    should be called as:
\begin{verbatim}
pincode-client prepare "my pass phrase"
\end{verbatim}
   This will generate a file \lstinline{verifkey.hex} that can be used
   as an argument when deploying the \lstinline{PubkeyRecovery} smart
   contract, and a file \lstinline{provkey.hex} that can be used to
   create new witnesses;
 \item To create a witness that the passphrase is known when providing
   a new public key. Because the witness contains the new public key,
   an attacker cannot intercept the message and replace it. The client
   should be called as:
\begin{verbatim}
pincode-client prove "my pass phrase" "PUBKEY_AS_HEX_NUMBER"
\end{verbatim}
  The command expects to find the file \lstinline{provkey.hex} and
  generates a file \lstinline{proof.hex} that should be submitted to
  the \lstinline{PubkeyRecovery} smart contract together with the new
  public key.
  \end{itemize}
\item[PubkeyRecovery smart contract:] This smart contract is a
  contract à la TIP-3 contract, using the initial public key to verify
  its address. This contract is very simple. It contains:
  \begin{itemize}
  \item A constructor to set the circuit that will be used to verify
    that the passphrase is known when it is used;
  \item An external function
    \lstinline{SetFromPincode(bytes proof, uint256 pubkey)}
    to define the new public key, while providing a
    proof of knowledge of the passphrase;
  \item An external function
    \lstinline{RecoverPubkey(RecoverablePubkey addr)} to call a
    recoverable contract to update the corresponding pubkey.
  \end{itemize}
\item[Abstract RecoverablePubkey contract:] This is an abstract
  contract, from which other contracts should inherit to be able to
  benefit from the {\tt PubkeyRecovery} mechanism. The contract
  provides two main functions:
  \begin{itemize}
  \item An external function \lstinline{SetPubkeyRecoveryCode(TvmCellcode)} to define the code of the \lstinline{PubkeyRecovery} smart
    contract. Because the code hash is known (inlined in the code),
    anybody can call this function with the correct code.
  \item An external function \lstinline{RecoverPubkey(uint256 oldkey,uint256 newkey)} that can only be called by a
    \lstinline{PubkeyRecovery} smart contract. The function verifies
    that the address belongs to such a contract before calling a
    function \lstinline{recover\_pubkey(oldkey, newkey)}, that the
    inheriting contract should define to specify the business logic
    that should happen when a pubkey is replaced.
  \end{itemize}
\item[Modified Multisig Wallet:] The
  \lstinline{RecoverableMultisigWallet} contract is the standard
  multisig wallet of Surf, updated for version 0.40 of Solidity, and
  modified to inherit from the \lstinline{RecoverablePubkey}
  contract. It defines the business logic to replace a custodian key.
  For that, it first checks that the old key belongs to an existing
  custodian of the contract, that the new key does not belong to an
  existing custodian, and then replace the old key by the new one,
  associated to the same index (for bitfields).
\end{description}

>>>>

\begin{description}
	\item[Code:] All our code comes with a README, standard
          in-code documentation and tests. This document is provided
          as an overview of our use cases for \zksnarks{} on the
          Freeton blockchain but you can also directly dive into the
          code. All our code is under the GPL license and is hosted at
          \url{https://github.com/OCamlPro/devex-18-zk-contest}.

@@

	\item[Keys on contracts: ] In the Euler and Pincode smart
          contracts, we provided, along with the verification key as
          needed by the \verb|vergroth16| instruction, the proving key
          (in a zipped format) on the smart contract. This enables
          users to easily find it in order to generate their
          proofs. In case the contract deployer should decide to keep
          the possibility of changing the verification and proving
          keys of the problem(s), this also makes it easiest to follow
          changes before submitting a solution. Unfortunately, the
          proving key for the Sudoku has size 64KB (19KB after
          compression) which is about 6 times (respectively 10 times
          once zipped) bigger than for Euler and Pincode, and we
          decided not to include it.

@@

\section{A toy example: the Sudoku grid}
\label{section_sudoku}

@@

\emph{Note: This section may be skipped upon a first reading without
  loss of information.}

@@

	\begin{itemize}
	\item To create the original proving and verification keys. This step is independent of any given Sudoku instance and should only be done once.
              \begin{verbatim}
              ./sudoku-client --sudoku-generate-keys
              \end{verbatim}
	    \item To generate a proof given an instance and a solution.
              \begin{verbatim}
              ./sudoku-client --sudoku-generate-proof --instance
                instance.in --solution solution.in
              \end{verbatim}
              The instance is given as a text file such as

@@

		Note that if your solution is not correct, this step will fail.
	\end{itemize}
      \item[The Sudoku smart contract:] The only contract for this use
        case is the \verb|Sudoku| contract. Only the deployer's public
        key is allowed to submit new Sudoku instances. Its constructor
        defines an initial instance and the verification key (in this
        case the proving key is not included in the smart contract due
        to its size). The main function \verb|submit| takes as input a
        proof and returns a boolean stating its correctness. The local
        function \verb|pi_from_instance| encodes the primary input
        given the current instance, but does not store it
        anywhere. The \verb|submit_instance| allows the owner to add a
        new challenge at any time.

@@

The previous example is somewhat limited by the fact that Sudokus are
easily solvable problems, either by hand or with a computer; they are
not the most convincing use case of (at least the zero-knowledge
aspect of) \zksnarks{}\footnote{Although they hint at a possible use
  for \zksnarks{}: checking a result without implementing its full
  logic in Solidity may save precious storage and computations on the
  blockchain.}.

@@

@@

\begin{description}
	\item [The C++ client]
	The `euler-client` program can be used in 2 ways:
	\begin{enumerate}
		\item \begin{verbatim}
		./euler-client prepare PROBLEM NONCE SOLUTION
		\end{verbatim}, where \verb|PROBLEM| is
		a problem number, \verb|NONCE| is an arbitrary number different for every
		problem, and \verb|SOLUTION| is the numeric solution, will generate a
		proving key and a verification key for the problem and its solution. These files can be published for users to propose submissions and to verify them.

		\item \begin{verbatim}
		./euler-client prove PROBLEM NONCE SUBMISSION PUBKEY
		\end{verbatim}, where
		\verb|SUBMISSION| is the numeric solution proposed by the user and
		\verb|PUBKEY| is his own public key (to prevent other users from
		submitting the same solution). The program 	generates a proof to be then
		submitted online.
	\end{enumerate}

	\item[Solidity Smart Contracts] The Solidity smart contracts are located in the `contracts/`
	directory. They can be built with `make` if `ft` (freeton-wallet) is
	installed.

	The following smart contracts are defined:
        \begin{enumerate}

	\item \verb|EulerRoot| : it's the central contract of a set of
        Euler problems. The \verb|EulerRoot| contracts is owned by the
        organizer, who is the only one able to publish new Euler
        problems.  Its most important functions are
        \verb|new_problem()| to deploy a new problem contract,
        \verb|new_user()| to deploy a user contract (anybody can call
        it) and \verb|submit()| to submit a solution for a
        problem. \verb|problem_address()| and \verb|user_address()|
        are useful to get the contract address of a problem or a
        user. An event is emitted everytime a problem is solved by a
        user.

	\item \verb|EulerProblem|: it's the contract corresponding to a
        given Project Euler problem. The contract contains the
        verification key of the problem, so that it can check the
        proofs submitted by users. Since proofs are attached to users'
        pubkeys, other users cannot re-use proofs submitted by other
        users. Each \verb|EulerProblem| contract contains the top 10
        of the first users to submit a correct solution. An event is
        also emitted everytime the problem is solved by a user. The
        contract also contains the description of the problem (title,
        description, url) and the information needed to submit
        solutions (compressed proving key and nonce).

	\item \verb|EulerUser|: it's the contract corresponding to the
        pubkey of a user. The user can provide a name using the
        \verb|set_name()| function.  The \verb|has_solved()| function
        is called by the \verb|EulerRoot| contract when the user has
        solved a new problem. The function stores the problem and the
        time of solution in the contract.

	For internal messages between these contracts, they either check that
	messages are coming from the unique \verb|EulerRoot| contract, or that they
	come from pre-computed addresses à la TIP-3.

\end{description}

\subsection{Brute-force protection}
\label{subsec:brute-force}
\emph{This small section details an interesting aspect we had to
  develop to prevent bruteforce attacks.}

To prevent users from easily brute-forcing the numeric solutions, we
added a simple proof of work of about 20 seconds. The proof of work is
based on 1,000,000 iterations of sha256 over the concatenation of the
problem number, the nonce and the solution. Since a random nonce
included in the hash, the proof-of-work cannot be done before the new
problem has been submitted and the corresponding new nonce published.

%\begin{description}
%	\item[The cli:]
%	\item[The EulerRoot TIP3-style contract:] Following the TIP3 conventions, this is the root contract to two types of leaves: EulerUser and EulerProblem. The EulerUser contract represents one user and the list of problems she has solved. The EulerProblem contract represents one problem and contains, among other things, the top ten leaderboard of the fastest solvers.
%\end{description}

@@

This use case includes 3 different components:
\begin{description}
\item[PinCode Client:] it is a C++ program linked to the Blueprint
  \zksnarks{} library. It can be used in two different ways:
  \begin{itemize}
  \item To create the initial circuit using the passphrase. The client
    should be called as:
\begin{lstlisting}
pincode-client prepare "my pass phrase"
\end{lstlisting}
   This will generate a file \lstinline{verifkey.hex} that can be used
   as an argument when deploying the \lstinline{PubkeyRecovery} smart
   contract, and a file \lstinline{provkey.hex} that can be used to
   create new witnesses;
 \item To create a witness that the passphrase is known when providing
   a new public key. Because the witness contains the new public key,
   an attacker cannot intercept the message and replace it. The client
   should be called as:
\begin{verbatim}
pincode-client prove "my pass phrase" "PUBKEY_AS_HEX_NUMBER"
\end{verbatim}
  The command expects to find the file \lstinline{provkey.hex} and
  generates a file \lstinline{proof.hex} that should be submitted to
  the \lstinline{PubkeyRecovery} smart contract together with the new
  public key.
  \end{itemize}
\item[PubkeyRecovery smart contract:] This smart contract is a
  contract à la TIP-3 contract, using the initial public key to verify
  its address. This contract is very simple. It contains:
  \begin{itemize}
  \item A constructor to set the circuit that will be used to verify
    that the passphrase is known when it is used;
  \item An external function
    \lstinline{SetFromPincode(bytes proof, uint256 pubkey)}
    to define the new public key, while providing a
    proof of knowledge of the passphrase;
  \item An external function
    \lstinline{RecoverPubkey(RecoverablePubkey addr)} to call a
    recoverable contract to update the corresponding pubkey.
  \end{itemize}
\item[Abstract RecoverablePubkey contract:] This is an abstract
  contract, from which other contracts should inherit to be able to
  benefit from the {\tt PubkeyRecovery} mechanism. The contract
  provides two main functions:
  \begin{itemize}
  \item An external function \lstinline{SetPubkeyRecoveryCode(TvmCellcode)} to define the code of the \lstinline{PubkeyRecovery} smart
    contract. Because the code hash is known (inlined in the code),
    anybody can call this function with the correct code.
  \item An external function \lstinline{RecoverPubkey(uint256 oldkey,uint256 newkey)} that can only be called by a
    \lstinline{PubkeyRecovery} smart contract. The function verifies
    that the address belongs to such a contract before calling a
    function \lstinline{recover\_pubkey(oldkey, newkey)}, that the
    inheriting contract should define to specify the business logic
    that should happen when a pubkey is replaced.
  \end{itemize}
\item[Modified Multisig Wallet:] The
  \lstinline{RecoverableMultisigWallet} contract is the standard
  multisig wallet of Surf, updated for version 0.40 of Solidity, and
  modified to inherit from the \lstinline{RecoverablePubkey}
  contract. It defines the business logic to replace a custodian key.
  For that, it first checks that the old key belongs to an existing
  custodian of the contract, that the new key does not belong to an
  existing custodian, and then replace the old key by the new one,
  associated to the same index (for bitfields).
\end{description}
